Species richness destabilizes trophic communities in freshwater ecosystems

### 2020-01-16 manuscript/bef_stability/manuscript.Rmd
Add predict and Rsq for figure one

```{r}
myload(temporal_community_metrics, dir = mypath("data"))

@@

stab_div_troph <- temporal_network_metrics %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med)
```

@@

st_basin <- get_basin_station()
myload(nmds, habitat_pressure, dir = mypath("report"))
myload(op_analysis, dir = mypath("data"))

@@

```{r mod-div-total}
ctrl <- lmeControl(opt='optim')
stab_div_mod <- lme(data = sem_data,
  fixed = log_stab ~ log_rich,
  random = ~ 1 + log_rich | basin,
  control = ctrl 
)
#stab_div_colin <- car::vif(stab_div_mod)
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke")
```

```{r predict}
pred <- ggpredict(stab_div_mod, terms = c("log_rich"), type = "fe")
stab_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_stab = predicted,
    log_rich = x,
    basin = group
  )

@@

```{r s-div, fig.dim = c(10,5)}
label_rsq <- "paste(
  \"Conditional italic(R) ^ 2 == \", round(stab_div_sum$Conditional, 2),
  \", Marginal italic(R) ^ 2 == \", round(stab_div_sum$Marginal, 2))"
# Plot left  
p_stab_div <- sem_data %>%
  ggplot(aes(x = log_rich, y = log_stab))+
  geom_point() +
  #scale_x_log10() + scale_y_log10() +
  labs(x = "Median species richness (Log 10)", y = " Biomass stability (Log 10)") +
  geom_line(data = stab_div_pred) +
  geom_ribbon(data = stab_div_pred, aes(ymin = conf.low, ymax = conf.high), fill
  = "grey20", alpha = .5) +
  annotate("text", x = 1.0, y = .8, label = paste(
  " Marg. R^2 = ", round(stab_div_sum$Marginal, 2),
  ", Cond. R^2 = ", round(stab_div_sum$Conditional, 2)
  )
    )

# Right plot
p_stab_div_troph <- stab_div_troph %>%
  ggplot(aes(x = richness_med, y = biomass_stab, color = troph_group)) +
  geom_point() +
  geom_smooth(method = "lm" ) +
  ggplot2::scale_color_manual(
    labels = troph_group_labeller(),
    values = c("blue", "green", "red"),
    name = "Trophic group"
  ) +
  scale_x_log10() + scale_y_log10() +
  labs(x = "Average species richness", y = "Biomass stability")

>>>>

```{r}
myload(temporal_community_metrics, dir = mypath("data"))
st_basin <- get_basin_station(sf_obj = FALSE)

@@

stab_div_troph <- temporal_network_metrics %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_stab = log10(biomass_stab)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, log_rich, log_stab) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

myload(nmds, habitat_pressure, dir = mypath("report"))
myload(op_analysis, dir = mypath("data"))

@@

with $i$ being the station $i$.

```{r mod-div-total}
ctrl <- lmeControl(opt='optim')
stab_div_mod <- lme(data = sem_data,
  fixed = log_stab ~ log_rich,
  random = ~ 1 + log_rich | basin,
  control = ctrl 
)
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred <- ggpredict(stab_div_mod, terms = c("log_rich"), type = "fe")
stab_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_stab = predicted,
    log_rich = x,
    basin = group
  )
```

```{r mod-div-troph}
stab_troph_mod <- lme(data = na.omit(stab_div_troph),
  fixed = log_stab ~ log_rich * troph_group,
  random = ~ 1 + log_rich * troph_group | basin,
  control = ctrl 
)

#stab_div_colin <- car::vif(stab_troph_mod) not use if one response
stab_troph_sum <- piecewiseSEM::rsquared(stab_troph_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred_troph_mod <- ggpredict(stab_troph_mod, terms = c("log_rich", "troph_group"), type = "fe")
stab_troph_pred <- pred_troph_mod %>%
  as_tibble() %>%
  rename(
    log_stab = predicted,
    log_rich = x,
    troph_group = group
  )

@@

```{r s-div, fig.dim = c(10,5)}
# Plot left  
p_stab_div <- sem_data %>%
  ggplot(aes(x = log_rich, y = log_stab))+
  geom_point() +
  #scale_x_log10() + scale_y_log10() +
  labs(x = "Median species richness (Log 10)", y = " Biomass stability (Log 10)") +
  geom_line(data = stab_div_pred) +
  geom_ribbon(data = stab_div_pred, aes(ymin = conf.low, ymax = conf.high), fill
  = "grey20", alpha = .5) +
  annotate("text", x = 1.0, y = .8, label = make_label_rsq(rsq = stab_div_sum),
    parse = TRUE)

# Right plot
p_stab_div_troph <- stab_div_troph %>%
  ggplot(aes(x = log_rich, y = log_stab, color = troph_group)) +
  geom_point() +
  ggplot2::scale_color_manual(
    labels = troph_group_labeller(),
    values = c("blue", "green", "red"),
    name = "Trophic group"
  ) +
  geom_line(data = stab_troph_pred) +
  geom_ribbon(data = stab_troph_pred,
    aes(ymin = conf.low, ymax = conf.high),
    alpha = .5) +
  annotate("text", x = 0.8, y = .8, label = make_label_rsq(rsq = stab_troph_sum),
    parse = TRUE) +
  #scale_x_log10() + scale_y_log10() +
  labs(x = "Median species richness", y = "Biomass stability")
